Analysing the three-dimensional structure of {\em P. falciparum}

### 2017-07-24 doc/manuscript/paper.tex
Finished the ensemble methods section

@@

@@

Both consensus and ensemble methods have benefits and drawbacks. Ensemble
approaches are more biologically accurate: Hi-C data are derived from a
population of cells, each of those with a uniquely folded 3D structure. A
population of structure thus better may represents the variability of
structures among ours cells. Yet, in addition of being more complex and
computationally intensive, ensemble approaches raise the question of
interpretability. Often, one has to fall back to interpreting the mean
structure \citep{kalhor:genome}, or a reduced set of structures
\citep{rousseau:three}. On the other hand, consensus approaches yield a single
structure recapitulating the rich information provided by Hi-C data, including
hallmarks of genome architecture shared across all cells despite the
cell-to-cell variability \citet{nagano:single-cell}. More amenable for
visualization and analysis, this average structure can be easily integrated
with other sources of data, such as RNA-seq, chip-seq etc, which are also
population-based.

@@

A typical 3C/Hi-C experiment can be summarized as a contact
count matrix $\mathbf{C} \in R^{n \times n}$, where each row and column
correspond to a genomic-window (or loci), and each entry $c_{ij}$ to the
number of times loci $i$ and $j$ have been observed interacting. Contact count
matrices are subjected to biological and technical biases that need to be
corrected \citep{imakaev:iterative, cournac:normalization, yaffe:probabilistic,
hu:hicnorm}.

Methods model chromosomes as a series of beads in 3D, each bead corresponding
to a specific genomic window. We denote by $\textbf{X} \in \RR^{3 \times n}$
the coordinate matrix of the structure(s) and $x_i \in \RR^3$ corresponds to
the coordinate matrix of the i-th beads.

@@

The first step is thus to find an adequate count-to-distance mapping to
convert contact counts into pairwise wish-distances $\delta_{ij}$.
\citet{dekker:capturing} models the 78 pairwise contact count of {\em S.
cerevisiae}'s chromosome III as a Worm-like chain polymer,
\citet{duan:three-dimensional} uses a linear mapping between contact counts
and wish distances, \citet{ay:three-dimensional} relies on the biophysical
properties of fractal globule polymers, \citet{tanizawa:mapping} inferred the
count-to-distance mapping by fitting the relationship using known pairwise
distances obtained through high-resolution FISH measures. While parameters of
these counts-to-distance functions vary, most of them can be summarized as a
power-law relationship : $c_{ij} = \beta d_{ij}^\alpha$.

\citet{lesne:3d} proposes a different approach to convert counts into
distances: they borrow the concept of shortest path from graph theory. A
graph is constructed from the contact map by considering each loci as a node.
Loci seen interacting are connected through an edge of weight
$\frac{1}{c_{ij}}$. The authors then compute wish-distances $\delta_{ij}$ as
the shortest path between node $i$ and $j$ in the graph. Constructing
wish-distances in such a way has two advantages: (i) low contact counts do not
contribute much to the wish-distances; (ii) the resulting wish-distances form
a distance matrix, and thus an optimal solution to the MDS can be found.

@@

\begin{table*}
\scriptsize
\centering
\begin{tabular}{cccccc}
\hline
\textbf{Paper} & Organism & Full or partial genome & Counts-to-distance mapping & \textbf{Constraints}
& \textbf{Weights}\\
\hline
\cite{dekker:capturing} & {\em S. cerevisiae} & Chr III & Worm-like chain behavior & - \\
\cite{duan:three-dimensional} & {\em S. cerevisiae} & Whole-genome & Linear
relationship & Adjacent beads, rDNA & - \\
\cite{tanizawa:mapping} & {\em S. pombe} & Whole-genome &
FISH-derived relationship & Adjacent beads, rDNA, telomeres, centromeres & - \\
\cite{ay:three-dimensional} & {\em P. falciparum} & Whole-genome & Fractal
globule derived relationship & adjacent
beads & $w_{ij} = \frac{1}{\delta_{ij}^2}$ \\
\cite{lesne:3d} & - & Whole-genome & \textit{ad hoc} & - & - \
\end{tabular}
\caption{Differences between MDS-based methods}{}
\label{table:mds_detail}
\end{table*}

@@

A crucial step of MDS-based method is the conversion of counts into
wish-distances. As described earlier, converting contact counts into
wish-distances require strong assumptions that may not be met in practice. For
example, this mapping may change from one organism to another
\citep{fudenberg:higher-order}, from one resolution to another
\citep{zhang:inference}, from one genomic distance range to another
\citep{ay:statistical}, or from one time point to another during the cell
cycle \citep{le:high-resolution, ay:three-dimensional}.

Two methods have since been proposed to alleviate this problem.
\citet{ben-elazar:spatial} proposes to cast the inference as a
\emph{non-metric MDS} \citep{kruskal:multidimensional}, where the 3D structure
is inferred jointly with the wish-distances. \fixme{except they do a bunch of
stuff to the count data}.

@@

A series of methods cast the inference as an optimization method, where the
objective function is designed by hand to fullfill a number of properties
\citep{trieu:large, trieu:3D, trieu:MOGEN}.

For example, \citep{trieu:large} formulates an optimization problem that
attempts to place the beads such that, if a significant contact is observed
between $i$ and $j$, then $\| x_i - x_j \| < d_c$. On the other hand, if a
no significant contact is observed, then the two beads should be no closer
than $d_c$. In addition, the authors also place constraints on the minimum
distance separating two beads and maximum distances between two beads. To
satisfy all those properties, the authors propose to minimize the following
objective function:

\fixme{add damn objective function}

This method is then generalized to include \textit{trans} contact counts data
\citep{trieu:MOGEN}.

\cite{trieu:MOGEN} proposes to exploit the Lozentzian function, to formulate a
constraint-based problem. They first convert the contact counts $c_{ij}$ into
wish-distances, and cast the following optimization problem:

@@

\citet{varoquaux:statistical} proposes a novel approach based on
statistical modeling of contact counts as random Poisson variables where the
intensity is a function of the distance: $c_{ij} \sim \text{Poisson}(\beta
d_{ij}^\alpha)$, and some perform the optimization only including a subset of
contact counts. The authors propose to cast the
inference of 3D models as maximizing the likelihood:

@@

@@

\citet{bau:three-dimensional}
\cite{umbarger:three-dimensional}
\cite{hu:bayesian}

@@

The last category of data-driven methods to infer the 3D architecture of the
genome rely on a new protocol to probe single-cell for their 3D structures
\citep{nagano:single-cell, XXX}. Single-cell Hi-C is still in its early days,
and only a handful of data sets are today publicly available. The contact maps
originating from these data sets are very sparse, and specific methods to
infer 3D structures need to be developed specifically for sc-HiC. The
challenge of constructing models from this data has been tackled very
differently by two groups.

\citet{nagano:single-cell} models the chromosome as a series of beads linked
as a flexible fiber. Loci seen interacting in the single-cell contact maps are
constrained to be in contact. The authors then sample a population of possible
structures.

\citet{paulsen:manifold} formulate a \textit{manifold based optimization},
where a low rank PSD matrix (and thus a distance matrix) is optimized to be as
close as possible to the sparse contact count matrix. Applying classical MDS
on this low rank PSD matrix then yields a 3D model of the genome.

\begin{table}[ht!]
\begin{center}
\footnotesize
\begin{tabular}{rlccccc}
\hline
\multirow{2}{*}{\emph{\footnotesize Publication}} & \multirow{2}{*}{\emph{\footnotesize Name}} &
\multirow{2}{*}{\emph{\footnotesize Con.}} &
\multirow{2}{*}{\emph{\footnotesize Pop.}}
&\multirow{2}{*}{\emph{\footnotesize MDS-based}} & \emph{\footnotesize Statistical} &
\multirow{2}{*}{\emph{\footnotesize Availability}} \\
& & & & & \emph{\footnotesize model} & \\
\hline
\hline
\footnotesize{\cite{dekker:capturing}} & & & & & & \\
\footnotesize{\cite{duan:three-dimensional}} &  & x & & x & & x \\
\footnotesize{\cite{tanizawa:mapping}} & & x & & x & & \\
\footnotesize{\cite{ay:three-dimensional}} & & x & & x & & \\
\footnotesize{\cite{ben-elazar:spatial}} & & x & & x & & x \\
\footnotesize{\cite{varoquaux:statistical}} & Pastis & x & & & x & x\\
\footnotesize{\cite{bau:three-dimensional}} & & & x & & & \\
\footnotesize{\cite{umbarger:three-dimensional}} & & x & & & &\\
\footnotesize{\cite{zhang:inference}} & chromSDE & x & & x & & x\\
\footnotesize{\cite{rousseau:three}} & & & x & & x & x\\
\footnotesize{\cite{hu:bayesian}} & Bach & x & x & & x & x\\
\footnotesize{\cite{kalhor:genome}} & & & x & & &\\
\footnotesize{\cite{lesne:3d}} & ShRec3D & x & & x & & x \\
\footnotesize{\cite{trieu:large}} & & x & & & & \\
\footnotesize{\cite{trieu:3D}} & & x & & & & \\
\footnotesize{\cite{trieu:MOGEN}} & MOGEN & x & & & & \\
\footnotesize{\cite{nagano:single-cell}} & & x & & & & \\
\footnotesize{\cite{paulsen:manifold}} & & x & & x & & x \\
\hline
\end{tabular}
\end{center}
\caption{\bf A comparison of 3D inference methods}{In this table, we summarize
properties of published methods to infer the 3D structure of the genome: (1)
is it a consensus or a population based inference? (2) Is it an MDS based
method? (3) or relies on a statistical modeling; (4) is the software
available or not (to the best of our knowledge).}
\end{table}

@@

The budding yeast \textit{S. cerivisiae}'s 3D structure has been extensively
studied, both through 3C-type studies \citep{dekker:capturing,
duan:three-dimensional, burton:} and through bio-imaging experiments
\citep{XX:}.

\citet{tjong:physical} showed \textit{S. cerevisiae}'s structure can be mostly
explained by a large set of constrained yet random population of structures
following a simple volume exclusion model. We repeated the experiment on {\em
P. falciparum}, and showed that, despite recovering similar results on {\em S.
cerevisiae}, the same set of constraints is not sufficient to explain {\em P.
falciparum}'s Hi-C contact map.

Briefly, let us recall the methodology presented by \citet{tjong:physical}. We
construct a population of three-dimensional structures by modeling chromosomes
as random configuration of beads-on-a-string, subjected to the following constraints.

\begin{itemize}
\item \textbf{Adjacent Beads Constraints} Two adjacent beads spaced 3.2kb apart are constrained to be 30nm apatr.
\item \textbf{Volume Exclusion Constraints} Each pairs of beads are spaced at minimum XX apart
\item \textbf{Nucleus Constraint} All beads lie in a spherical nucleus of specific
radius (XXX)
\item \textbf{Centromeres Constraints} Centromeres are localized in a small sphere of
50nm abutting the nuclear envelop.
\item \textbf{Telomere Constraints} Telomeres are located at most 50nm from the nuclear envelop.
\end{itemize}

\citet{tjong:physical} then formulates an optimization problem that includes
in addition to all the aforementioned constraints a penalty term on three
adjacent beads accounting for chromatin stiffness.

\begin{equation}
\frac{1}{2} k_{\text{angle}} \sum^{N - 2}_{i = 1} \left( 1 - \frac{x_{i + 1} -
x_i}{\|x_{i + 1} - x_i\|} \cdot \frac{x_{i + 2} - x_{i + 1}}{\|x_{i + 2} -
  x_{i + 1}\|} \right)^2\,,
\end{equation}

\citet{tokuda:dynamical}

\citet{wong:how} model the budding yeast chromosomes a semi-flexible fiber
constrained in a nucleus, and applies 4 sequences specific forces on this
fiber to obtain certain properties: (1) centromeres are attached to a single
point of the nucleus by a segment; (2) telomeres are subjected to an outward
force, that pushes them towards the nuclear membrane; (3) rDNA is thicken; (4)
apply a random brownian movement. This model recovers many of the known
hallmarks of the 3D architecture of {\em S. cerevisiae}.

>>>>

\usepackage{booktabs}
\usepackage[svgnames,table]{xcolor}
\usepackage[tableposition=above]{caption}
\usepackage{pifont}

\newcommand*\CHECK{\ding{51}}

@@

\todo{A revoir}

@@

Both consensus and ensemble methods have benefits and drawbacks. Ensemble
approaches are more biologically accurate: Hi-C data are derived from a
population of cells, each of those with a uniquely folded 3D structure. The
variability of structures among ours cells may thus be better represented by a
population of structures. Yet, in addition of being more complex and
computationally intensive, ensemble approaches raise the question of
interpretability. Often, one has to fall back to interpreting the mean
structure \citep{kalhor:genome}, or a reduced set of structures
\citep{rousseau:three}. On the other hand, consensus approaches yield a single
structure recapitulating the rich information provided by Hi-C data, including
hallmarks of genome architecture shared across all cells despite the
cell-to-cell variability \citet{nagano:single-cell}. More amenable for
visualization and analysis, this average structure can be easily integrated
with other sources of data, such as RNA-seq, chip-seq etc, which are also
population-based.

@@

A typical 3C/Hi-C experiment can be summarized as a contact
count matrix $\mathbf{C} \in \RR^{n \times n}$, where each row and column
correspond to a genomic-window (or loci), and each entry $c_{ij}$ to the
number of times loci $i$ and $j$ have been observed interacting. Contact count
matrices are subjected to biological and technical biases that need to be
corrected \citep{imakaev:iterative, cournac:normalization, yaffe:probabilistic,
hu:hicnorm}.

Methods model chromosomes as a series of beads in 3D, each bead corresponding
to a specific genomic window. We denote by $\textbf{X} \in \RR^{3 \times n}$
the coordinate matrix of the structure(s), $x_i \in \RR^3$ corresponds to
the coordinate matrix of the i-th beads and $d_{ij} = \|x_i - x_j\|_2$ the
euclidean distance separating beads $i$ and $j$.

@@

The first step is thus to find an adequate count-to-distance mapping to
convert contact counts into pairwise wish-distances $\delta_{ij}$.
\citet{dekker:capturing} models the 78 pairwise contact count of {\em S.
cerevisiae}'s chromosome III as a Worm-like chain polymer,
\citet{duan:three-dimensional} uses a linear mapping between contact counts
and wish distances, \citet{ay:three-dimensional} relies on the biophysical
properties of fractal globule polymers, \citet{tanizawa:mapping} inferred the
count-to-distance mapping by fitting the relationship using known pairwise
distances obtained through high-resolution FISH measures. While parameters of
these counts-to-distance functions vary, most of them can be summarized as a
power-law relationship: $c_{ij} = \beta d_{ij}^\alpha$.

\citet{lesne:3d} proposes a different approach to convert counts into
distances: they borrow the concept of shortest path from graph theory to
create a distance matrix. A graph is constructed from the contact map by
considering each loci as a node. Loci seen interacting are connected through
an edge of weight $\frac{1}{c_{ij}}$. The authors then compute wish-distances
$\delta_{ij}$ as the shortest path between node $i$ and $j$ in the graph.
Constructing wish-distances in such a way has two advantages: (i) low contact
counts do not contribute much to the wish-distances; (ii) the resulting
wish-distances form a distance matrix, and thus an optimal solution to the MDS
can be found.

@@

\begin{table*}
\scriptsize
\centering
\begin{tabular}{rccccccccc}
\hline
\multirow{2}{*}{\textbf{Paper}} & \multirow{2}{*}{\textbf{Organism}} &
\textbf{Full or}
& \textbf{Counts-to-distance} &
\multicolumn{4}{c}{\multirow{2}{*}{\textbf{Constraints}}}
& \multirow{2}{*}{\textbf{Weights}}\\
 & & \textbf{partial} & \textbf{mapping} &  \\
 \cmidrule(lr){5-8} 
 & & & & Adj. & rDNA & C & T & \\
\hline
\cite{dekker:capturing} & {\em S. cerevisiae} & Chr III & Worm-like chain
behavior & & & & \\
\cite{duan:three-dimensional} & {\em S. cerevisiae} & Whole-genome & Linear
relationship & \CHECK &  \CHECK & \CHECK &  & \\
\cite{tanizawa:mapping} & {\em S. pombe} & Whole-genome &
FISH-derived relationship & \CHECK & \CHECK & \CHECK & \CHECK &  \\
\cite{ay:three-dimensional} & {\em P. falciparum} & Whole-genome & Fractal
globule derived relationship & \CHECK  & & &
 & $\frac{1}{\delta_{ij}^2}$ \\
\cite{lesne:3d} & & Whole-genome & \textit{ad hoc} & & & & & \
\end{tabular}
\caption{Differences between MDS-based methods}{}
\label{table:mds_detail}
\end{table*}

@@

A crucial step of MDS-based method is the conversion of counts into
wish-distances. As described earlier, converting contact counts into
wish-distances require strong assumptions that may not be met in practice. For
example, this mapping changes from one organism to another
\citep{fudenberg:higher-order}, from one resolution to another
\citep{zhang:inference} or even from one time point to another during the cell
cycle \citep{le:high-resolution, ay:three-dimensional}. Several methods have
since been proposed to alleviate this problem.

\citet{ben-elazar:spatial} casts the inference as a \emph{non-metric MDS}
\citep{kruskal:multidimensional}, where the 3D structure is inferred jointly
with the wish-distances. The authors first filter the
interaction counts to keep only the most significant interactions and
interpolate the missing values to obtain a smooth, symmetric positive definite
matrix.

@@

A series of methods cast the inference as an optimization method, where the
objective function is designed by hand to fulfill a number of properties
\citep{trieu:large, trieu:MOGEN, trieu:3D}. For example, \citep{trieu:large}
formulates an optimization problem that attempts to place the beads such that,
if a significant contact is observed between $i$ and $j$, then $\| x_i - x_j
\| < d_c$. On the other hand, if a no significant contact is observed, then
the two beads should be no closer than $d_c$. In addition, the authors also
place constraints on the minimum distance separating two beads and maximum
distances between two beads. To satisfy all those properties, the authors
propose to minimize the following objective function:

\begin{equation*}
\begin{aligned}
\mathcal{O}(X, \delta_{ij}) = & \underset{i, j | c_{ij} \neq 0, |i-j| \neq 1}{\sum} \Bigg(W_1 \text{tanh}(d^2_c -
d^2_{ij}) c_{ij} + W_2 \frac{\text{tanh}(d^2_{ij} - d^2_\text{min})}{N} \Bigg) \\
& \underset{i, j | c_{ij} = 0}{\sum} \Bigg(W_3 \frac{\text{tanh}(d^2_\text{max} -
d^2_{ij})}{N} + W_4 \frac{\text{tanh}(d^2_{ij} -
d^2_{c})}{N}\Bigg) \\
& \underset{i, j | |i - j| = 1}{\sum} \Bigg(W_1 \frac{\text{tanh}(d^2_\text{amax} -
d^2_{ij})}{N} + W_2 \frac{\text{tanh}(d^2_{ij} -
d^2_\text{min})}{N}\Bigg) \,,\\
\end{aligned}
\end{equation*}

where $N$ is the total number of interaction, $(W_1, W_2, W_3, W_4)$ weights
associated to each term, $d_\text{amax}$ the maximum distance separating two
adjacent beads, $d_\text{min}$ the minimum distance separating two adjacent
beads and $d_\text{max}$ the maximum distance between two beads. 
This method is then generalized to include \textit{trans} contact counts data
\citep{trieu:MOGEN}

To give another example of carefully crafted optimization, \cite{trieu:3D}
proposes to exploit the Lozentzian function, to formulate a constraint-based
problem. They first convert the contact counts $c_{ij}$ into wish-distances,
and cast the following optimization problem:

@@

\citet{varoquaux:statistical} proposes a novel approach called \texttt{Pastis}
based on statistical modeling of contact counts as random Poisson variables
where the intensity is a function of the distance: $c_{ij} \sim
\text{Poisson}(\beta d_{ij}^\alpha)$, and some perform the optimization only
including a subset of contact counts. The authors propose to cast the
inference of 3D models as maximizing the likelihood:

@@

\texttt{Pastis} can automatically adjust the parameters $\beta$, $\alpha$ of
the counts-to-distance transfer function and infer a genome structure that
best explains the observed data.

@@

{\em Ensemble approaches} aim at inferring a population of structures
representative of the contact count map. The methods fall into two distinct
categories: the first type of problem casts a non-convex or ill-defined
optimization problem and sample local minima of the function
\citep{bau:three-dimensional, umbarger:three-dimensional}, while the second
type proposes a statistical modeling of the problem and samples the posterior
distribution \citep{rousseau:three, hu:bayesian}.A parallel can be made
between the first group of methods and {\em Consensus} MDS type of approaches,
and the second with {\em Consensus} statistical based methods.

\subsubsection*{Sampling local minima}

\citet{umbarger:three-dimensional,bau:three-dimensional, kalhor:genome} models
chromosomes as a series of beads, linked by restraining oscillators. These
oscillators can be thought of as a ``force'' between beads so that they come
into contact or ensure a minimal or maximal distance between those. The models
includes three differente types of restraints: (i) beads seen interacting are
restrained with harmonic oscillators of strengths derived from the contact
counts; (ii) adjacent beads are ensured to be neither too close nor too far
from one another; (iii) the last oscillators are activated only when
constraints are not fulfilled to XXX.  This yields an optimization problem
with a large number of local minima, which the authors sample from by running
50,000 minimizations starting from random starts.

\subsubsection*{Estimating the posterior distribution}

Similarly to \citet{varoquaux:statistical}, \citet{rousseau:three} and
\citet{hu:bayesian} proposes to model contact counts with a formal
probabilistic model. \citet{rousseau:three} models observed contact counts
$c_{ij}$ as a random Gaussian variable of mean $\beta d_{ij}^{\alpha}$,
$\alpha \leq 0$ and variance $\sigma_{ij}$ estimated directly from the contact
count data, while \citet{hu:bayesian} models contact counts as random Poisson
variables of mean $\beta d_{ij}^\alpha$. The authors then sample from the
posterior using MCMC.

@@

The last category of data-driven methods to infer the 3D architecture of the
genome rely on a new protocol to probe single-cell for their 3D structures
\citep{nagano:single-cell,ramani:massively}. Single-cell Hi-C is still in its early days,
and only a handful of data sets are today publicly available. The contact maps
originating from these data sets are very sparse, and specific methods to
infer 3D structures need to be developed specifically for sc-HiC. The
challenge of constructing models from this data has been tackled very
differently by two groups.

Akin to {\em consensus} MDS-like method and {\em ensemble}-local minima
methods,  the first approach is to consider each contact as a constraint and
to formulate
an under constrained optimization problem. A population of structures
satistifying the constraints can be found by sampling local minimas.

\citet{paulsen:manifold} proposes a very different method, and formulate a
\textit{manifold based optimization}, where a low rank PSD matrix (and thus a
distance matrix) is optimized to be as close as possible to the sparse contact
count matrix. Applying classical MDS on this low rank PSD matrix then yields a
3D model of the genome.

\begin{table}[ht!]
\begin{center}
\footnotesize
\begin{tabular}{rlccccc}
\hline
\multirow{2}{*}{\textbf{\footnotesize Publication}} & \multirow{2}{*}{\textbf{\footnotesize Name}} &
\textbf{\footnotesize Consensus or} 
&\multirow{2}{*}{\textbf{\footnotesize MDS-based}} & \textbf{\footnotesize Statistical} &
\multirow{2}{*}{\textbf{\footnotesize Available}} \\
& & \textbf{\footnotesize Ensemble}
 & & \textbf{\footnotesize model} & \\
\hline
\hline
\footnotesize{\cite{dekker:capturing}} & & C & & & \\
\footnotesize{\cite{duan:three-dimensional}} &  & C & \CHECK & & \CHECK \\
\footnotesize{\cite{tanizawa:mapping}} & & C & \CHECK & & \\
\footnotesize{\cite{ay:three-dimensional}} & & C & \CHECK & & \\
\footnotesize{\cite{ben-elazar:spatial}} & & C & \CHECK & & \CHECK \\
\footnotesize{\cite{varoquaux:statistical}} & Pastis & C & & \CHECK & \CHECK\\
\footnotesize{\cite{bau:three-dimensional}} & & E & & &  \\
\footnotesize{\cite{umbarger:three-dimensional}} & & E  & & &\\
\footnotesize{\cite{zhang:inference}} & chromSDE & C &  \CHECK & & \CHECK\\
\footnotesize{\cite{rousseau:three}} & & E & & \CHECK & \CHECK\\
\footnotesize{\cite{hu:bayesian}} & Bach & E/C &  & \CHECK & \CHECK\\
\footnotesize{\cite{kalhor:genome}} & & E &   & &\\
\footnotesize{\cite{lesne:3d}} & ShRec3D & C  & \CHECK & & \CHECK \\
\footnotesize{\cite{trieu:large}} & & C & & & \\
\footnotesize{\cite{trieu:3D}} & & C & &  & \\
\footnotesize{\cite{trieu:MOGEN}} & MOGEN & C & & & \\
\footnotesize{\cite{nagano:single-cell}} & & E & & & \\
\footnotesize{\cite{paulsen:manifold}} & & C & \CHECK & & \CHECK \\
\hline
\end{tabular}
\end{center}
\caption{\bf A comparison of 3D inference methods}{In this table, we summarize
properties of published methods to infer the 3D structure of the genome: (1)
is it a consensus or a ensemble based inference? (2) Is it an MDS based
method? (3) or relies on a statistical modeling; (4) is the software
available or not (to the best of our knowledge).}
\end{table}

@@

The budding yeast \textit{S. cerivisiae}'s 3D structure has been extensively
studied, both through 3C-type studies \citep{dekker:capturing,
duan:three-dimensional, burton:} and through bio-imaging experiments
\citep{XX}. The small size of its genome, the well-known hallmarks of its
genome architecture and the availability of high resolution contact maps and
FISH data set quickly led several teams to investigate the minimal set of
constraints needed to reproduce the hallmarks of its genome architecture.

\cite{tjong:physical, tokuda:dynamical, tjong:physical} model {\em S.
cerevisiae}'s chromosomes as flexible random fibers under a small set of
constraints. While the exact modeling proposed by the three groups differ, the
set of constraints are similar: \textit{(i)} the chromosomes are constrained
into a spherical ball representing the nucleus; \textit{(ii)} centromeres are
constrained into a spherical ball tethered to the nuclear membrane;
\textit{(iii)} telomeres are tethered to the nuclear membraine; \textit{(iv)}
rDNA is constrained into the nucleolus, represented as a spherical ball
opposite to the centromeres.

Let us take a closer look at the model proposed by \citet{tjong:physical}.
The 16 chromosomes are modeled as beads-on-a-string spaced 3.2kb apart. The
authors cast an optimization problem including the following restraints:

\begin{itemize}
\item \textbf{Adjacent Beads Constraints} Two adjacent beads spaced 3.2kb
apart are constrained to be 30~nm apart;
\item \textbf{Volume Exclusion Constraints} Each pairs of beads are spaced at
minimum $30~nm$ apart;
\item \textbf{Nucleus Constraint} All beads lie in a spherical nucleus of
radius $1\mu m$
\item \textbf{Centromeres Constraints} Centromeres are localized in a small sphere of
50nm abutting the nuclear envelop.
\item \textbf{Telomere Constraints} Telomeres are located at most 50nm from the nuclear envelop.
\item \textbf{rDNA restraint} The repeated rDNA located on chromosome 12 is
constrained to be inside the nucleolus;
\item \textbf{Nucleolus exclusion restraint} All beads that isn't rDNA is
constrained outside of the nucleolus;
\item \textbf{Chain stiffness}: an angular restraint is imposed on three
consecutive beads to reproduce chain stiffness:
\begin{equation}
\frac{1}{2} k_{\text{angle}} \sum^{N - 2}_{i = 1} \left( 1 - \frac{x_{i + 1} -
x_i}{\|x_{i + 1} - x_i\|} \cdot \frac{x_{i + 2} - x_{i + 1}}{\|x_{i + 2} -
  x_{i + 1}\|} \right)^2\,,
\end{equation}
\end{itemize}

In the end, the objective function minimized is the following:

\begin{equation}
\end{equation}

Using IMP, the authors generate a 25,000 structures, all of which respect the
constraints listed above. The population of structures is then used to created
a ``volume-exclusion contact map'', considering that two beads that are less
then 45~nm apart from a contact. The Pearson correlation of the
volume-exclusion contact map and the Hi-C one are highly correlated,
demonstrating this small set of constraints fully explain the observed counts.
In addition, the population of structures also explain FISH experiments
previously published.
